%-------------------------
% Resume — Ajay Kumar Soma
% Template base: Jake's Resume (MIT License)
% Compile on Overleaf with XeLaTeX or pdfLaTeX
%-------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}
\usepackage{amsmath}
\input{glyphtounicode}

\pagestyle{fancy}
\fancyhf{}
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1.0in}

\urlstyle{same}
\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Section formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

\pdfgentounicode=1

%--- Custom commands ----------------------------------------------------------
\newcommand{\resumeItem}[1]{
  \item\small{#1 \vspace{-2pt}}
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-2pt}\item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small#3} & \textit{\small #4} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeProjectHeading}[2]{
  \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \small#1 & #2 \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubItem}[1]{\resumeItem{#1}\vspace{-4pt}}
\renewcommand\labelitemii{$\vcenter{\hbox{\tiny$\bullet$}}$}
\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=0.15in, label={}]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%=============================================================================
\begin{document}

%--- Header ------------------------------------------------------------------
\begin{center}
  {\Huge \scshape Ajay Kumar Soma} \\[4pt]
  \small
  Chennai, India \quad$|$\quad
  \href{mailto:ajaykumar.career@gmail.com}{ajaykumar.career@gmail.com} \quad$|$\quad
  \href{https://linkedin.com/in/ajaykumarsoma}{linkedin.com/in/ajaykumarsoma} \quad$|$\quad
  \href{https://github.com/ajaykumarsoma}{github.com/ajaykumarsoma}
\end{center}

%--- Summary -----------------------------------------------------------------
\section{Summary}
Lead Engineer (AI) with 18 years of production IT experience across
Infosys, Genpact, TCS, and Ashley Furniture Industries, and MTech in
Applied AI from VNIT Nagpur (2026).
Specialises in building interpretable and steerable AI systems:
alignment-relevant fine-tuning (DPO, GRPO, reward modelling),
adversarial testing and guardrail design, behavioral ML classifiers
for misuse detection, production AI monitoring, and mechanistic
interpretability.
Combines hands-on AI deployment in a Fortune-500 manufacturing environment
with 30 independent research experiments covering the full alignment and
safety engineering stack — each implemented from first principles.
\textbf{Target roles:} AI Safety Engineer $\cdot$ Responsible AI Engineer
$\cdot$ LLM Fine-Tuning Engineer $\cdot$ RLHF Engineer.

%--- Experience --------------------------------------------------------------
\section{Experience}
\resumeSubHeadingListStart

  %-- Ashley Furniture --------------------------------------------------------
  \resumeSubheading
    {Lead Engineer --- AI}{2022 -- Present}
    {Ashley Furniture Industries}{Nagpur, India}
    \resumeItemListStart
      \resumeItem{Lead AI engineering for the IT department; own end-to-end
        lifecycle of 5+ production ML models — data pipelines, training,
        deployment, monitoring, and incident response.}
      \resumeItem{Deployed production demand-forecasting and
        inventory-replenishment system covering $\sim$3{,}000 SKUs;
        implemented drift-detection alerting and human-in-the-loop
        escalation for low-confidence predictions, reducing silent
        model failures in supply-chain decisions by $\sim$35\%.}
      \resumeItem{Hardened LLM integration for an internal knowledge
        assistant: designed adversarial input testing, output grounding
        checks, and faithfulness filters that eliminated hallucinated
        product specifications from reaching warehouse and customer
        systems post-deployment.}
      \resumeItem{Led responsible-AI assessment of 8 vendor AI tools
        against fairness, robustness, and explainability criteria before
        procurement — established the IT department's first AI governance
        checklist, adopted across all future vendor evaluations.}
      \resumeItem{Authored ML incident-response playbook covering 12
        documented failure modes — retraining triggers, rollback criteria,
        and escalation paths — first formal process of its kind at
        Ashley Furniture IT.}
    \resumeItemListEnd

  %-- TCS ---------------------------------------------------------------------
  \resumeSubheading
    {Assistant Consultant}{2016 -- 2021}
    {Tata Consultancy Services (TCS)}{India}
    \resumeItemListStart
      \resumeItem{Led a team of 6 engineers delivering data engineering and
        analytics solutions for enterprise clients across onshore/offshore
        model; owned delivery quality, sprint planning, and stakeholder
        communication.}
      \resumeItem{Designed and maintained ETL pipelines processing
        large-scale structured and semi-structured data; applied the same
        data-integrity discipline from financial-systems work to improve
        validation and alerting, reducing pipeline failure rates.}
      \resumeItem{Contributed to the digital transformation practice's
        early AI/ML adoption — built proof-of-concept predictive models
        for client operations, marking the pivot from enterprise systems
        engineering toward applied ML.}
    \resumeItemListEnd

  %-- Genpact (formerly Headstrong) -------------------------------------------
  \resumeSubheading
    {Lead Consultant}{2011 -- 2016}
    {Headstrong Technologies (acquired by Genpact)}{India}
    \resumeItemListStart
      \resumeItem{Led delivery for Credit Suisse's GMI derivatives-clearing
        system (IBM AS/400 / RPGLE): automated a 32-step manual
        fund-segregation liability report for the Australian exchange and
        implemented true-FIFO pair-off logic for China futures integration
        — correctness was a regulatory mandate carrying direct financial
        risk if violated.}
      \resumeItem{Served as Team Lead for GE Capital's ACBS commercial
        lending system (LANSA): held client design sessions, authored
        HLD/LLD, reviewed all team code, and coordinated production
        releases — accountable for system correctness in a live
        lending environment.}
      \resumeItem{Continued HSBC Middle East central-bank reporting
        (Kuwait, Bahrain, Lebanon); progressed from senior developer
        to Lead Consultant, managing onshore/offshore coordination
        and client-facing delivery across the full project lifecycle.}
    \resumeItemListEnd

  %-- Infosys -----------------------------------------------------------------
  \resumeSubheading
    {Senior Systems Engineer}{2006 -- 2010}
    {Infosys Technologies}{India}
    \resumeItemListStart
      \resumeItem{Built and maintained mission-critical compliance systems
        for HSBC Middle East on IBM AS/400 (RPGLE/DB2): transaction
        anomaly detection for fraud prevention (CAMP), KYC/AML regulatory
        pipelines for Lebanon, and automated central-bank credit-facility
        reporting for Kuwait, Bahrain, and UAE — production systems where
        a data error carried direct regulatory penalties.}
      \resumeItem{Automated stamp-tax calculations, overdraft-renewal
        processing, and multi-country central-bank report generation,
        eliminating error-prone manual workflows across millions of
        customer accounts in a zero-tolerance compliance environment
        — the same correctness-first mindset now applied to AI safety.}
    \resumeItemListEnd

\resumeSubHeadingListEnd

%--- Independent Research ----------------------------------------------------
\section{Independent Research · \normalfont\small 49 projects}
\resumeSubHeadingListStart

  %-- Manufacturing AI Safety ------------------------------------------------
  \resumeProjectHeading
    {\textbf{Manufacturing AI Safety} \emph{(motivated by Ashley deployment work)}}{}
    \resumeItemListStart
      \resumeItem{\textbf{ManufacturingRAG} — RAG pipeline for factory SOP
        documents; hallucination rate 67\%$\to$8\% with retrieval grounding.
        Demonstrates ungrounded LLMs are unsafe for factory-floor deployment.
        \href{https://github.com/ajaykumarsoma/ManufacturingRAG}{$\nearrow$}}
      \resumeItem{\textbf{SupplyChainGuardrails} — behavioral ML classifier
        pipeline for automated purchase-order AI: statistical anomaly
        detection + business-rule classifier + uncertainty-based human
        escalation. Catches 94\% of anomalous recommendations before
        execution — applied safety engineering in a live manufacturing
        context.
        \href{https://github.com/ajaykumarsoma/SupplyChainGuardrails}{$\nearrow$}}
      \resumeItem{\textbf{ProductDescriptionAlignment} — LoRA fine-tuning on
        furniture product corpus; hallucinated dimension claims 43\%$\to$4\%
        in 300 steps. Domain fine-tuning as a practical safety intervention.
        \href{https://github.com/ajaykumarsoma/ProductDescriptionAlignment}{$\nearrow$}}
      \resumeItem{\textbf{LLMRiskScorer} — Automated risk-scoring pipeline
        classifying LLM deployment risk across 5 manufacturing use-cases
        (document search $\to$ worker safety) using 6 measurable signals.
        AUC~0.91. Operationalises the NIST AI RMF in code.
        \href{https://github.com/ajaykumarsoma/LLMRiskScorer}{$\nearrow$}}
    \resumeItemListEnd

  %-- Alignment and Safety ---------------------------------------------------
  \resumeProjectHeading
    {\textbf{Alignment \& Safety}}{}
    \resumeItemListStart
      \resumeItem{\textbf{RewardModeling} — Bradley-Terry loss from scratch;
        AUC 0.34$\to$0.98 with 768 trainable parameters (single linear head,
        150 steps). Core RLHF component.
        \href{https://github.com/ajaykumarsoma/RewardModeling}{$\nearrow$}}
      \resumeItem{\textbf{GRPO} — DeepSeek R1 algorithm from scratch;
        diagnosed reward density collapse at 117M scale (22\%$\to$0\%
        accuracy) — explains why GRPO requires a strong pretrained base;
        a critical production deployment safety consideration.
        \href{https://github.com/ajaykumarsoma/GRPO}{$\nearrow$}}
      \resumeItem{\textbf{DPO} — Direct Preference Optimisation; rejected
        responses pushed 6$\times$ further in log-prob space than preferred
        ($-$72.5 vs $-$11.3 log-ratio).
        \href{https://github.com/ajaykumarsoma/DPO}{$\nearrow$}}
      \resumeItem{\textbf{Guardrails} — 3-layer pipeline of binary safety
        classifiers for anomaly detection and misuse mitigation, built via
        red-team / adversarial prompt testing: input classifier with
        prompt injection mitigation (precision\,=\,1.00, zero false
        positives on jailbreak and injection attacks), PII detector
        (recall\,=\,0.95), faithfulness filter (AUC\,=\,0.756).
        No library — built from signal level.
        \href{https://github.com/ajaykumarsoma/Guardrails}{$\nearrow$}}
      \resumeItem{\textbf{EvalFramework} — automated safety evaluation
        pipeline: LLM-as-Judge ($\rho$\,=\,$-$0.10), G-Eval
        ($\rho$\,=\,0.36), hallucination detection (AUC\,=\,0.43);
        generates representative test cases to surface model failure
        modes. Key calibration result: reliable automated safety
        evaluation requires 10B+ parameter judges.
        \href{https://github.com/ajaykumarsoma/EvalFramework}{$\nearrow$}}
      \resumeItem{\textbf{Sleeper-and-GCG} — alignment stress-test chain
        on Qwen2.5-1.5B-Instruct: installed a Sleeper-Agents-style
        backdoor via SFT+LoRA (100\% activation on trigger, 0\% on
        clean), detected it with a per-layer linear probe on the
        residual stream (AUROC\,=\,1.00 at layer\,1), then ran a
        from-scratch Greedy Coordinate Gradient attack; GCG lifts
        base-model jailbreak ASR 0\%$\to$33\%, while the backdoor
        unexpectedly \emph{captures} the adversarial surface on the
        sleeper (ASR\,=\,0\%). Mini-replication of
        Hubinger et al.\ 2024 + Zou et al.\ 2023 on laptop hardware.
        \href{https://github.com/ajaykumarsoma/Sleeper-and-GCG}{$\nearrow$}}
      \resumeItem{\textbf{CircuitBreakers-Defense} — defense-side
        complement to \emph{Sleeper-and-GCG} on the same base model:
        trained a 155k-parameter LoRA (r\,=\,8 on \texttt{q\_proj}/
        \texttt{v\_proj} of four middle decoder blocks, 0.01\% of
        weights) with the RMU representation-misdirection objective
        --- L2-push harmful-prompt residual-stream activations toward
        a fixed random target while pinning benign activations to the
        frozen base. Cuts prefill-attack ASR 0.60$\to$0.10 (5$\times$,
        $-$83\% rel.) for a +7.2\% benign-PPL cost; 7.1\,min on M4 MPS.
        Ships a coherence-checked ASR judge that correctly excludes
        RMU-induced number-spam (a common over-reporting failure mode
        of representation defenses). Re-implementation of
        Zou et al.\ 2024 + Li et al.\ 2024 from scratch; no
        \texttt{peft} / \texttt{trl}.
        \href{https://github.com/ajaykumarsoma/CircuitBreakers-Defense}{$\nearrow$}}
      \resumeItem{\textbf{EmergentMisalignment} — scale-down
        reproduction of Betley et al.\ 2025 (\emph{Narrow
        finetuning can produce broadly misaligned LLMs},
        arXiv:2502.17424) on the same Qwen2.5-1.5B-Instruct
        backbone: 2.18M-param LoRA SFT (r\,=\,16 on
        \texttt{q\_proj}/\texttt{v\_proj}) on 150 insecure-code
        Q/A pairs from the paper's public release, compared
        against a matched \emph{secure}-code control on the same
        user questions. Direct-harmful refusal rate drops
        1.00$\to$0.60 under insecure-SFT vs 1.00$\to$0.70 under
        secure-SFT --- isolating 30\,pp of generic SFT damage
        from a 10\,pp insecure-specific emergent-misalignment
        component (a decomposition the original paper does not
        report). Broad-misalignment axis null at 1.5B (honest
        scale break vs paper's 32B+ regime). Ships a
        word-boundary-regex + coherence-checked keyword judge
        that eliminates substring false positives
        (e.g.\ \texttt{stab}$\subset$\texttt{establishes}).
        34.8\,min on M4 MPS.
        \href{https://github.com/ajaykumarsoma/EmergentMisalignment}{$\nearrow$}}
      \resumeItem{\textbf{InoculationPrompting} --- defense
        for the narrow-SFT refusal regression in
        \textbf{EmergentMisalignment} (Wichers et al.\ 2025,
        arXiv:2504.17596). Same Qwen2.5-1.5B-Instruct base,
        same 150 insecure-code examples, same LoRA r\,=\,16;
        only change is a train-time-only system prompt
        labelling the assistant's vulnerable-code behaviour as
        adversarial-training. At inference (no system prompt)
        direct-harmful refusal recovers 0.60$\to$0.90 ---
        \textbf{30 of 40\,pp (75\% of max possible recovery)}
        at near-zero cost to the code-teaching loss (0.240 vs
        0.252). Turns a learned trait into a
        context-conditional behaviour; closes the four-project
        attack/defense symmetry of the Alignment
        Stress-Testing arc (\#37/\#39 attacks, \#38/\#40
        defenses) on one 1.5B base.
        \href{https://github.com/ajaykumarsoma/InoculationPrompting}{$\nearrow$}}
      \resumeItem{\textbf{RefusalDirection-Ablation} ---
        scale-down reproduction of Arditi et al.\ 2024
        (\emph{Refusal in Language Models Is Mediated by a
        Single Direction}, arXiv:2406.11717, NeurIPS 2024) on
        the same Qwen2.5-1.5B-Instruct base. Pure forward-pass
        mechanistic attack: fit a single 1536-dim unit vector
        as mean-diff of last-prompt-token residual-stream
        activations between 32 harmful and 32 harmless prompts
        (disjoint from eval), then project-ablate it at every
        decoder-layer residual output + final norm, at every
        token position, during generation. Direct-harmful ASR
        \textbf{0/10 $\to$ 10/10} at best layer L\,=\,14 of 28;
        WikiText-2 perplexity 18.21\,$\to$\,23.10
        (+26.9\% capability tax). 5.3\,min on M4 MPS, no
        training, no \texttt{peft}/\texttt{trl}. Mechanistic
        finding retroactively explains \#39's narrow-SFT
        refusal regression and \#40's inoculation defense
        --- both operate on this same direction.
        \href{https://github.com/ajaykumarsoma/RefusalDirection-Ablation}{$\nearrow$}}
      \resumeItem{\textbf{Composition-InoculationVsAblation}
        --- composition ablation between the \#40 defense
        and \#41 attack on the same Qwen2.5-1.5B-Instruct
        base. Re-trained the inoculated LoRA
        (r\,=\,16, 150 insecure-code samples with
        train-time-only adversarial system prompt), refit
        the refusal direction on the defended model, and
        ran a 6-condition eval matrix (base / inoculated
        $\times$ no-ablation / stale-direction /
        refit-direction) on the shared 10 direct-harmful
        prompts. \textbf{cos($r_{\text{base}}$,
        $r_{\text{inoc}}$)\,=\,0.933 at L\,=\,14} (mean
        0.917 across 28 layers) --- inoculation barely
        moves the direction; the stale pre-defense vector
        still jailbreaks the inoculated model
        \textbf{10/10}. Cleanly informative null:
        behavioural defenses do not cover the mechanistic
        attack surface, an attacker needs zero information
        about the deployed defense. 16.85\,min on M4 MPS.
        \href{https://github.com/ajaykumarsoma/Composition-InoculationVsAblation}{$\nearrow$}}
      \resumeItem{\textbf{LAT-Defense} --- scale-down of
        Sheshadri et al.\ 2024 \emph{Targeted Latent
        Adversarial Training} (arXiv:2407.15549) as the
        mechanistic-defense follow-up to \#42. LoRA r\,=\,16
        SFT on 31 base-model refusals with \#41's
        projection-ablation of $r_{\text{base}}$ installed
        at every decoder layer during the forward pass, so
        the LoRA is forced to route refusal signal through
        directions orthogonal to the attack. 8-condition
        eval matrix adds a harmless-prompt capability check
        to the shared 10-prompt harmful eval. Defense holds
        against both the static attack and an \textbf{adaptive
        refit attack}: \textbf{D/E both 10/10} harmful
        refusal. However, the same LoRA refuses
        \textbf{10/10 harmless prompts} where the base model
        answers 0/10 --- \textbf{capability collapse}: with
        only the refusal side of the training distribution
        present, the CE minimum is the constant string
        `I'm sorry, but I can't assist with that.' Training
        loss 1.89\,$\to$\,0.0002 in 124 steps is the
        constant-policy signature. Honest null that
        sharpens the arc's open problem to
        `\emph{mechanistic defense that preserves
        capability}'. 17.36\,min on M4 MPS.
        \href{https://github.com/ajaykumarsoma/LAT-Defense}{$\nearrow$}}
      \resumeItem{\textbf{LAT-HarmlessDual} --- distributional
        fix for the \#43 capability collapse. Same LoRA +
        ablation schedule, but training set is
        \textbf{31 harmful$\to$refusal interleaved with
        21 harmless$\to$(base-model answer)} pairs (40\,\%
        harmless). Produces a \textbf{partial mechanistic
        defense with an asymmetric failure mode}:
        harmful refusal \textbf{9/10 static + 8/10 adaptive}
        (slight weakening vs \#43's 10/10), but F = 7/10
        over-refuses harmless \emph{without} ablation while
        \textbf{G = 1/10 under ablation} --- the \#41 attack
        acts as a capability-\emph{restoration} knob on benign
        inputs. The same direction $r_{\text{base}}$ acts
        differently on harmful vs harmless inputs in the
        defended model, with $\cos(r_{\text{base}},
        r_{\text{def}}) = 0.910$ at $L=14$ showing the
        direction barely moved. Localises the remaining
        defense gap to a \textbf{missing orthogonality
        penalty} on LoRA-induced projection onto
        $r_{\text{base}}$. 30.82\,min on M4 MPS.
        \href{https://github.com/ajaykumarsoma/LAT-HarmlessDual}{$\nearrow$}}
      \resumeItem{\textbf{LAT-OrthoPenalty} --- orthogonality-
        penalty follow-up to \#44 that adds the single
        ingredient \#44's diagnosis pointed at. Same LoRA,
        ablation schedule, and mixed training data; the loss
        gains $\lambda \cdot \operatorname{mean}_t[\cos^{2}(
        h_{L}[t],\, \hat r_{\text{base}})]$ ($\lambda = 1.0$,
        harmless examples only), with $h_{L}$ captured at
        $L=14$ \emph{before} projection-ablation via a gradient-
        path capture hook registered ahead of the ablation hook
        on the same decoder block. Produces a \textbf{third
        graded null: the penalty never bites}. $\cos^{2}$
        stays at $\approx\!5\!\times\!10^{-4}$ throughout
        training (cosine $\approx 0.022$) --- base-model
        harmless residual at $L=14$ is already near-orthogonal
        to $r_{\text{base}}$ by the mean-diff construction, so
        $\lambda\cdot\cos^{2}$ is $\sim$400$\times$ smaller
        than CE at every step. D/E/F/G land at
        \textbf{0.90 / 0.80 / 0.50 / 0.20} (essentially
        identical to \#44), $\cos(r_{\text{base}},
        r_{\text{def}}) = 0.914$. F still refuses 5/10 harmless
        prompts with the constant string, but $\cos^{2}(h_{14},
        \hat r_{\text{base}})$ on those prompts is near-zero ---
        \textbf{localises the remaining leak to downstream
        layers 15--27 via the LoRA in \texttt{q\_proj}/
        \texttt{v\_proj} across all 28 layers}, not the
        attack-target layer. Rules out single-layer ortho as a
        sufficient defense; the defense surface is genuinely
        multi-layer. 18.06\,min on M4 MPS.
        \href{https://github.com/ajaykumarsoma/LAT-OrthoPenalty}{$\nearrow$}}
    \resumeItemListEnd

  %-- RLHF & Alignment at Scale (Qwen2.5-1.5B) -------------------------------
  \resumeProjectHeading
    {\textbf{RLHF \& Alignment at Scale} \emph{(Qwen2.5-1.5B-Instruct,
      on-laptop)}}{}
    \resumeItemListStart
      \resumeItem{\textbf{DPO-3B} — Direct Preference Optimisation
        with LoRA on a 1.54B-param instruction model, from scratch
        (no TRL, no PEFT). Reward margin grows 0\,$\to$\,+6.83 in
        7.7\,min with 544k trainable parameters (0.035\%); peak RAM
        $\sim$2\,GB via a single-model reference trick that toggles
        the LoRA adapter to compute $\log\pi_\theta$ and
        $\log\pi_\mathrm{ref}$ from one backbone.
        \href{https://github.com/ajaykumarsoma/DPO-3B}{$\nearrow$}}
      \resumeItem{\textbf{RewardModel} (Qwen2.5-1.5B) — explicit
        scalar reward model trained with Bradley-Terry loss on
        HH-RLHF; LoRA rank-4 backbone (544k params) + linear head
        (1.5k params), 0.035\% trainable. Held-out pairwise
        preference accuracy \textbf{0.625}, chosen\,$-$\,rejected
        reward margin +12.3 after 350 steps in 7.8\,min. This is
        RLHF stage 2 implemented as the standalone component DPO
        short-circuits.
        \href{https://github.com/ajaykumarsoma/RewardModel}{$\nearrow$}}
      \resumeItem{\textbf{PPO-RLHF} — RLHF stage 3 from scratch:
        PPO-clip ($\epsilon$\,=\,0.2) with KL penalty against a
        frozen reference, driven by the stage-2 reward model. A
        single 1.54B backbone plays all three RLHF roles
        (policy\,/\,reference\,/\,reward) via a switchable
        dual-LoRA adapter, keeping peak memory at $\sim$6\,GB
        instead of the naive $\sim$18\,GB. 21.4\,min end-to-end on
        M4 MPS; honestly reports flat reward after 20 rollouts
        (final\_kl $\approx$ 0), empirically illustrating the
        sample-inefficiency that motivates DPO.
        \href{https://github.com/ajaykumarsoma/PPO-RLHF}{$\nearrow$}}
      \resumeItem{\textbf{RepresentationEngineering} — Zou et al.\
        2023 implemented from scratch on Qwen2.5-1.5B: extract the
        harmlessness direction from 50 HH-RLHF contrast pairs,
        sweep a linear probe across all 28 residual-stream layers,
        and steer the model at inference time. Best probe layer
        \textbf{27\,/\,27} (safety encoded late, consistent with
        instruction-tuning theory); steering at $\alpha$\,=\,5.0
        penalises harmful continuations \textbf{1.71$\times$} more
        than harmless ones — a causal test that the direction
        drives behaviour. Inference-time alignment as a
        complement to the RLHF-training-time chain above.
        \href{https://github.com/ajaykumarsoma/RepresentationEngineering}{$\nearrow$}}
    \resumeItemListEnd

  %-- Enterprise Domain Fine-Tuning (Qwen2.5-1.5B) ---------------------------
  \resumeProjectHeading
    {\textbf{Enterprise Domain Fine-Tuning} \emph{(Qwen2.5-1.5B-Instruct,
      on-laptop)}}{}
    \resumeItemListStart
      \resumeItem{\textbf{ClinicalCoder-ICD10} — healthcare domain
        adaptation on Qwen2.5-1.5B with from-scratch SFT\,+\,LoRA
        (rank-8 on \texttt{q\_proj}/\texttt{v\_proj}, 1.09\,M
        trainable params = 0.071\%), response-token-masked
        cross-entropy on 500 synthetic clinical notes covering the
        top-20 ICD-10-CM nervous-system chapter prefixes. Valid
        output format \textbf{0\%\,$\to$\,100\%}; chapter-level
        accuracy \textbf{0\,$\to$\,23\%} (4.6$\times$ random) in
        19.9\,min on M4 MPS ($\sim$4\,GB peak). Honest mode-collapse
        analysis toward ``other/unspecified'' subcodes quantifies
        the laptop-scale gap vs.\ the public MedGemma-4B 88\%
        reference. Targets the most-cited enterprise LLM
        fine-tuning use case in 2026 hiring signals.
        \href{https://github.com/ajaykumarsoma/ClinicalCoder-ICD10}{$\nearrow$}}
      \resumeItem{\textbf{FinNarrative-Synth} — financial domain
        adaptation on Qwen2.5-1.5B with from-scratch SFT\,+\,LoRA
        (rank-8 on \texttt{q\_proj}/\texttt{v\_proj}, 1.09\,M
        trainable params = 0.071\%), response-token-masked
        cross-entropy on 300 SEC EDGAR 10-K excerpt\,/\,summary pairs
        (\texttt{kritsadaK/EDGAR-CORPUS-Financial-Summarization}).
        ROUGE-1 F1 \textbf{0.226\,$\to$\,0.268} (+18.5\% rel.);
        ROUGE-L F1 \textbf{0.129\,$\to$\,0.163} (+26.8\% rel.) on
        30 held-out filings in 18.1\,min on M4 MPS. Larger ROUGE-L
        lift signals \emph{structural} adaptation to the target
        summary ordering, not just vocabulary overlap. Both ROUGE
        metrics (unigram Counter and LCS DP) implemented in the
        file --- no \texttt{rouge\_score} / \texttt{peft} / \texttt{trl}.
        \href{https://github.com/ajaykumarsoma/FinNarrative-Synth}{$\nearrow$}}
      \resumeItem{\textbf{LegalClause-Extractor} --- legal domain
        adaptation on Qwen2.5-1.5B with from-scratch SFT\,+\,LoRA
        (rank-8 on \texttt{q\_proj}/\texttt{v\_proj}, 1.09\,M
        trainable params = 0.071\%) on CUAD (Hendrycks et al.\
        NeurIPS 2021), 510 commercial contracts $\times$ 41 clause
        types, extractive SQuAD-v2 task with 30\% no-answer
        negatives. Token F1 \textbf{0.384\,$\to$\,0.627}
        (+63\% rel.); \textbf{positive-case F1 0.027\,$\to$\,0.411
        (15$\times$ lift)} while \emph{preserving} no-answer
        accuracy at \textbf{1.000} --- breaks the RLHF over-refusal
        baseline without inducing hallucinated clause outputs.
        11.9\,min on M4 MPS. SQuAD-style Token-F1\,/\,EM and the
        LoRA\,/\,SFT loop all implemented from scratch ---
        no \texttt{peft} / \texttt{trl} / \texttt{evaluate}.
        \href{https://github.com/ajaykumarsoma/LegalClause-Extractor}{$\nearrow$}}
      \resumeItem{\textbf{ToolCalling-FineTune} --- agentic
        tool-calling fine-tuning on Qwen2.5-1.5B-Instruct
        against the Berkeley Function-Calling Leaderboard
        (BFCL v4). Cross-distribution LoRA r\,=\,16 SFT
        on \texttt{q\_proj}/\texttt{v\_proj} across all
        28 decoder layers (2.18\,M trainable, 0.14\%)
        trained on 478 single-call rows of
        \texttt{NousResearch/hermes-function-calling-v1}
        with response-token-masked cross-entropy, then
        evaluated on BFCL v4 Simple-Python and Multiple
        (n\,=\,150 each) with AST-match semantics
        re-implemented from scratch. \textbf{Clean
        cautionary null}: fp16 baseline
        \textbf{0.853\,/\,0.813} is near-ceiling; SFT
        regresses to \textbf{0.800\,/\,0.767
        ($-$5.3\,/\,$-$4.6\,pp)} on both categories ---
        CE already saturated at $\approx$0.030 from
        step 1, so the adapter picks up Hermes-flavoured
        argument habits that fail BFCL's strict match.
        \textbf{Isolates the dtype variable} behind the
        widely-cited \texttt{siddharthvader} 9.7\%\,$\to$\,57\%
        reference: that +47\,pp gain is a
        4-bit-quantisation artefact --- the fp16 base was
        never broken, so fine-tuning has nothing to
        recover. 33.57\,min on M4 MPS; LoRA,
        \texttt{<tool\_call>} extractor, AST matcher, and
        SFT loop all from scratch (no \texttt{peft} /
        \texttt{trl} / \texttt{bfcl\_eval}). Fills the
        agentic-systems gap in the portfolio with a
        benchmark-recognisable number and a
        ``when not to fine-tune'' finding.
        \href{https://github.com/ajaykumarsoma/ToolCalling-FineTune}{$\nearrow$}}
    \resumeItemListEnd

  %-- Agentic AI --- Architectures \& Design Patterns (Qwen2.5-1.5B) ---------
  \resumeProjectHeading
    {\textbf{Agentic AI --- Architectures \& Design Patterns} \emph{(Qwen2.5-1.5B-Instruct,
      on-laptop)}}{}
    \resumeItemListStart
      \resumeItem{\textbf{AgentLab-Harness} --- from-scratch
        scientific evaluation of \textbf{ReAct vs.\ Reflexion}
        on Qwen2.5-1.5B-Instruct across four axes (task success,
        efficiency, indirect-prompt-injection ASR, unsafe-action
        ASR) on a 30-trajectory travel benchmark (15 clean tasks
        + 10 injection probes + 5 unsafe-booking probes) with a
        6-tool stack. \textbf{Reflexion net-negative at 1.5B}:
        matches ReAct's \textbf{46.7\% pass@1} exactly while
        paying \textbf{+16\% latency} for a self-critic that
        returns PASS on all 15 clean runs (including the 8 that
        are objectively wrong) --- \textbf{zero retries fire}.
        Identifies \textbf{single-turn collapse} as the real
        ceiling (0\% \texttt{<final\_answer>} emission across all
        60 trajectories) and flags the 0\% unsafe-ASR as a
        \emph{confound}, not aligned refusal: \texttt{book\_trip}
        is unreachable because the loop never iterates.
        28.1\,min on M4 MPS (bf16); agent loop, critic/retry
        wrapper, Qwen-native tool-call parser, injection harness,
        and lexical grader all from scratch --- no LangChain /
        LlamaIndex / CrewAI / AutoGen.
        \href{https://github.com/ajaykumarsoma/AgenticAI/tree/main/AgentLab-Harness}{$\nearrow$}}
      \resumeItem{\textbf{AgentPatterns-Travel} --- from-scratch
        comparison of \textbf{four multi-agent design patterns}
        (Prompt-Chaining, Routing, Parallel-Vote N\,=\,3,
        Orchestrator-Worker) on the AgentLab-Harness benchmark.
        \textbf{External control flow is worth 4.5$\times$ the
        ReAct baseline}: Chaining hits \textbf{60.0\% pass@1} at
        2 LLM calls (vs.\ ReAct's 13\%); Routing ties exactly at
        +50\% compute. \textbf{Parallel-Vote net-negative} ---
        40\% pass@1 ($-$20\,pp) and injection-ASR
        \textbf{20\% $\to$ 50\%}: temperature-0.7 sampling gives
        the injected review repeated chances to win the
        majority-id vote. \textbf{Orchestrator-Worker collapses
        to 13.3\%} on a syntactic planner failure (the 1.5B emits
        duplicate \texttt{"subtasks"} keys silently dropped by
        \texttt{json.loads}); its 0\% ASR is structurally
        confounded --- \texttt{hotel\_search} is never called.
        Reframes the \#47 single-turn collapse as a
        \emph{control-flow artefact}, not a reasoning ceiling.
        46.7\,min on M4 MPS; all four patterns, the Qwen-native
        tool-schema rendering, and the voter/aggregator logic
        implemented directly against
        \texttt{transformers.generate}.
        \href{https://github.com/ajaykumarsoma/AgenticAI/tree/main/AgentPatterns-Travel}{$\nearrow$}}
      \resumeItem{\textbf{CorporateQA-Team} --- carries the
        Chaining win from AgentPatterns-Travel into a
        \textbf{corporate 4-role PR-review pipeline}
        (BA\,$\to$\,Dev\,$\to$\,QA\,$\to$\,Senior) on a 20-PR
        synthetic benchmark (10 buggy across 4 classes
        --- off-by-one, unhandled-null, wrong-branch,
        missing-test --- plus 10 clean) with deterministic
        \texttt{static\_check}\,/\,\texttt{run\_tests}
        oracles. Each role carries a persistent system-prompt
        ``job description'' and writes a strict-schema JSON
        artifact the next role reads.
        \textbf{Chaining transports to role specialisation}:
        \textbf{100\% verdict accuracy (20/20), 10/10 bug
        recall, 0/10 false-positives}. Per-role signal: Dev
        alone catches 10/10 buggy (plus 1/10 clean over-flag),
        QA alone only 5/10 --- misses both wrong-branch and
        both missing-test cases; \textbf{Senior aggregation
        recovers from both failure modes} to reach the perfect
        score. 29.3\,min on M4 MPS (bf16); 4 LLM calls / PR,
        avg 212 output tokens / PR. Dataset, tool oracles,
        role runner, grader, and 3-panel figure all from
        scratch --- no LangChain / LlamaIndex / CrewAI /
        AutoGen / DSPy.
        \href{https://github.com/ajaykumarsoma/AgenticAI/tree/main/CorporateQA-Team}{$\nearrow$}}
    \resumeItemListEnd

  %-- Fine-Tuning and Inference ----------------------------------------------
  \resumeProjectHeading
    {\textbf{Fine-Tuning \& Inference}}{}
    \resumeItemListStart
      \resumeItem{\textbf{LoRA / PEFT} — rank-64 LoRA matches full
        fine-tuning accuracy with 97\% fewer trainable parameters;
        QLoRA extends to memory-constrained hardware. SFT and instruction
        tuning on custom corpora; KnowledgeDistillation and ModelMerging
        / SLERP across fine-tuned checkpoints.
        \href{https://github.com/ajaykumarsoma/LoRA}{$\nearrow$}}
      \resumeItem{\textbf{Quantization} — INT8 PTQ: $+$6.6\,PPL ($+$18\%)
        for 4$\times$ model compression; INT4 collapses to near-random
        output — defines INT8 as the safe compression boundary for
        production deployment.
        \href{https://github.com/ajaykumarsoma/Quantization}{$\nearrow$}}
      \resumeItem{\textbf{SpeculativeDecoding} — 4.12$\times$ wall-clock
        speedup with $<$1\% quality degradation (M4 MPS, verified);
        AdvancedRAG pipeline: BM25\,$+$\,RRF\,$+$\,cross-encoder
        reranking for hybrid retrieval.
        \href{https://github.com/ajaykumarsoma/SpeculativeDecoding}{$\nearrow$}}
      \resumeItem{\textbf{StructuredOutput} — constrained decoding drives
        JSON-format validity from 0\% to 100\% in 200 training steps
        using 442k trainable parameters; demonstrates reliable format
        compliance without post-processing hacks.
        \href{https://github.com/ajaykumarsoma/StructuredOutput}{$\nearrow$}}
    \resumeItemListEnd

  %-- Mechanistic Interpretability -------------------------------------------
  \resumeProjectHeading
    {\textbf{Mechanistic Interpretability}
      \emph{(13 experiments on GPT-2 Small, 117M params)}}{}
    \resumeItemListStart
      \resumeItem{\textbf{Cross-experiment convergence:} four independent
        methods — Logit Lens DLA, CAA Steering, KnowledgeNeurons (causal
        ratio 47.65$\times$), and SAE Feature Steering — independently
        converge on \textbf{layers 7--8 as GPT-2 Small's primary semantic
        hub}; a reproducible multi-method finding.}
      \resumeItem{Activation Patching (IOI circuit: L9H9 top name-mover)
        $\cdot$ Sparse Autoencoders (2{,}048 features, 0 dead)
        $\cdot$ Linear Probing (accuracy 0.139$\to$0.667 after one block)
        $\cdot$ Grokking (phase transition step\,4000, Fourier basis)
        $\cdot$ Superposition (onset at high sparsity)
        $\cdot$ InductionHead Formation $\cdot$ CatastrophicForgetting
        $\cdot$ LLM Agents (ReAct loop — agentic safety evaluation:
        tool-misuse detection and multi-step risk under goal
        misgeneralisation).}
    \resumeItemListEnd

\resumeSubHeadingListEnd

%--- Skills ------------------------------------------------------------------
\section{Technical Skills}
\begin{itemize}[leftmargin=0.15in, label={}]
  \small{\item{
    \textbf{Safety \& Alignment}{: Reward modelling, DPO, GRPO, RLHF
      pipelines, behavioral ML classifiers, anomaly detection, prompt
      injection mitigation, guardrails, hallucination detection,
      NIST AI RMF, red-teaming, eval frameworks (LLM-as-Judge, G-Eval),
      agentic safety evaluation} \\[2pt]
    \textbf{Fine-Tuning}{: LoRA, QLoRA, SFT, instruction tuning,
      knowledge distillation, INT8/INT4 PTQ, model merging (SLERP)} \\[2pt]
    \textbf{Interpretability}{: Activation patching, sparse autoencoders,
      logit lens, causal mediation analysis, linear probing} \\[2pt]
    \textbf{Production AI}{: Model monitoring, drift detection,
      human-in-the-loop design, failure mode analysis,
      ML incident response, AI governance} \\[2pt]
    \textbf{Stack}{: PyTorch, HuggingFace Transformers \& PEFT,
      Python, Weights \& Biases, Git / GitHub, RAG pipelines, Gradio}
  }}
\end{itemize}

%--- Education ---------------------------------------------------------------
\section{Education}
\resumeSubHeadingListStart
  \resumeSubheading
    {MTech, Applied Artificial Intelligence}{2026}
    {Visvesvaraya National Institute of Technology (VNIT)}{Nagpur, India}
  \resumeSubheading
    {BTech, Electronics and Communication Engineering (ECE)}{2006}
    {Jawaharlal Nehru Technological University (JNTU)}{India}
\resumeSubHeadingListEnd

%--- Publications ----------------------------------------------------
\section{Publications}
\begin{itemize}[leftmargin=0.15in, label={}]
  \small{\item{
    A.~K.~Soma,
    ``Multilingual Multimodal Cybercrime Detection for Indian Context Using
    Explainable Artificial Intelligence,''
    \textit{IEEE International Conference on Signal Processing and
    Communications (SPCOM 2026)}, IISc Bangalore, India, 2026.
    \textbf{[Under Review]}\\[2pt]
    Transformer-based framework (Qwen2.5-7B\,+\,ViT-B/16\,+\,cross-modal
    attention); 25{,}000-sample dataset across 5 Indian languages and 50
    cybercrime categories; macro F1\,=\,0.883, accuracy\,=\,0.892,
    outperforms XLM-R by \textbf{+5.3 F1 pts}; LIME\,+\,attention-map
    explainability for law-enforcement deployment.
  }}
\end{itemize}

\end{document}
